Bird beak's elaboration and exploration

### 2021-06-07 Analysis/03-elaboration_exploration_analyses.Rmd
changed var covar to var-covar plus added major axis figure

## Extracting the variance covariance matrices from the MCMCglmm mini-chains

@@

These major axes were calculated as the longest distance within the 95% confidence interval hyper-ellipse from the variance-covariance matrix (i.e. the major axes of the 95% CI ellipse of matrix). % NC: A small figure here would help
In practice, the coordinates of these axes end points were calculated directly from the cross product between the 95% unit hypersphere scaled by the variance-covariance matrix's squared eigenvalues and the variance-covariance matrix's eigenvectors.

## Major axes calculation
% NC: Mathematically seems sound, but I'm no expert. But very hard to parse for anyone who hasn't been looking at this in detail. Think fo the poor palaeontologists!

For $O_{n}$, the unit hypersphere matrix of _n_ dimensions and a radius composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

@@

Following the ways to: %NC: Not sure what you mean here
 1. estimate the variance-covariance matrices of the phylogenetic effects in trait space (02-MCMCglmm_mini_chains.Rmd);
 2. extracting the major axes of variation of these VCV matrices for the different levels in the phylogeny;
 3. and calculating the elaboration and exploration scores for elements in the trait space based on these major axes.

>>>>

## Extracting the variance-covariance matrices from the MCMCglmm mini-chains

@@

These major axes were calculated as the longest distance within the 95% confidence interval hyper-ellipse from the variance-covariance matrix (i.e. the major axes of the 95% CI ellipse of matrix). <!-- % NC: A small figure here would help -->

```{r, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5}
set.seed(42)
## Space plot
cor_matrix <- matrix(cbind(1, 0.8, 0.8, 1), nrow = 2)
space <- space.maker(50, 2, rnorm, cor.matrix = cor_matrix)
lim <- c(floor(range(space)[1]), ceiling(range(space)[2]))
plot(space, pch = 19, xlab = "Trait 1", ylab = "Trait 2", xlim = lim, ylim = lim)

## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix), col = "blue", lwd = 3)
## Plotting the major axis
lines(get.axes(list(list(list(VCV = cor_matrix, Sol = c(0,0)))))[[1]][[1]], lwd = 3, col = "orange")
```

Figure: the main axis of variation (in orange) of the 95% CI ellipse of the variance-covariance matrix (in blue) of 20 elements in 2 dimensions (in black).

In practice, the coordinates of these axes end points were calculated directly from the cross product between the 95% unit hypersphere scaled by the variance-covariance matrix's squared eigenvalues and the variance-covariance matrix's eigenvectors.

## Major axes calculation
<!-- % NC: Mathematically seems sound, but I'm no expert. But very hard to parse for anyone who hasn't been looking at this in detail. Think fo the poor palaeontologists!
 -->
For $O_{n}$, the unit hypersphere matrix of _n_ dimensions and a radius composed of the two identity matrices $I_{n}$ and $-I_{n}$ so that:

@@

Following the ways to: <!-- %NC: Not sure what you mean here -->
 1. estimate the variance-covariance matrices of the phylogenetic effects in trait space (02-MCMCglmm_mini_chains.Rmd);
 2. extracting the major axes of variation of these VCV matrices for the different levels in the phylogeny;
 3. and calculating the elaboration and exploration scores for elements in the trait space based on these major axes.
